VM-Centric Snapshot Deduplication\\ for Cloud Data Backup

### 2015-05-07 snapshot_system/mass_arch.tex
change for msst

We use the virtual device driver in Xen that employs a bitmap to track the changes 
that have been made to the virtual disk.
Every bit in the bitmap represents a fixed-sized (2MB) segment, indicating whether the segment
has been modified since last backup. 
Segments are further divided into variable-sized chunks (average 4KB) 
using a content-based chunking algorithm~\cite{frame05}, 
which brings the opportunity of fine-grained deduplication.
When the VM issues a disk write, the dirty bit for the corresponding segment is set
and this indicates such a segments needs to be checked during snapshot backup. 
After the snapshot backup is finished, the driver resets the dirty bit map to a clean state.
For data modification during backup, copy-on-write protection is set so that backup can continue to
copy  a specific version while new changes are recorded.
%copies a frozen version while the modification can still carry on.

The representation of each snapshot has  a two-level index data structure.
The snapshot meta data (called snapshot recipe) contains a list of segments, each of which contains segment
metadata of its chunks (called segment recipe).
In snapshot and segment recipes, 
the data structures  include references to the actual data location to eliminate the need for additional indirection.

@@

We use the QFS distributed file system to hold snapshot backups.
Following the VM-centric idea for the purpose of fault isolation,
each VM has its own snapshot store, containing new data chunks which are considered
to be non-duplicates.
As shown in Figure~\ref{fig:as_arch}, we explain the data structure of the snapshot stores as follows.
%\begin{itemize}
%\item
There is an independent store containing all PDS chunks shared among different VMs as
a single file.
%All PDS chunks are stored in one PDS file. 
Each reference to a PDS data chunk in the PDS index is the offset within the PDS file.
%There is an option to use a data structure similar to the non-PDS store. We opt
%for the simpler format because
Additional compression is not applied because 
for the data sets we have tested, we only observed limited spatial locality 
among popular data chunks. On average the number of consecutive PDS index hits is lower than 7,
thus it is not very effective to group a large number of chunks as a compression and data fetch unit. 
For the same reason, we decide not to take the sampled index approach~\cite{Guo2011} 
for detecting duplicates from PDS as limited spatial locality is not sufficient to enable
effective prefetching for sampled indexing.

@@

For non PDS data, the snapshot store of a VM is  divided into a set of containers and 
each container is approximately 1GB.  The reason for this  division is to simplify the compaction process
conducted periodically. When  chunks are deleted from old snapshots,
chunks without any reference from other snapshots can be removed by this compaction process.
By limiting the size of a container, we can control the length of each round of compaction.
The compaction  can work on one container at a time and move the in-use data chunks to another container.

Each non-PDS container is further divided into a set of chunk data groups. Each chunk group is composed of
a set of data chunks and is the basic unit in data access and retrieval. 
In writing a chunk during backup, the system accumulates data chunks and stores the entire
group as a unit after compression. This  compression can reduce data by several times  in our tested data.
When accessing a particular chunk, its chunk group is retrieved from the storage
and decompressed. Given the high spatial locality and usefulness of prefetching  in 
snapshot chunk accessing~\cite{Guo2011,Rhea2008},
retrieval of  a data chunk  group naturally works well with prefetching. 
A  typical chunk group contains 1000 chunks in our experiment.
% with an average size of 200-600 chunks.

%\item
Each non-PDS data container is represented by three files:
%in the DFS
1) the container data file holds the actual content, 
2) the container index file is responsible for translating a data reference
into its location within a container, and 
3) a chunk deletion log file records all the deletion requests within  the container.

%\item
A non-PDS data chunk reference stored in the index of snapshot recipes
is composed of two parts: a container ID with 2 bytes and a local chunk ID with 6 bytes.
Each container maintains a local  chunk counter and assigns the current number 
as a chunk ID  when  a new chunk is added to this  container. 
Since data chunks are always appended to a snapshot store during backup, 
local chunk IDs are monotonically increasing.
When a chunk is to be accessed, the segment recipe contains a reference pointing to  a data chunk, which is used to lookup up the chunk data as described shortly.

@@

\item
\noindent\textbf{Append()}. 
For PDS data, the chunk is appended to the end of the PDS file and the offset is returned as the  reference.
This operation  may only be used during PDS recalculation.
For non-PDS data, this call places a chunk into 
the snapshot store and returns a reference to be stored in the recipe of a snapshot. 
The write requests to append chunks to a VM store are accumulated at the client side. 
When the number of write requests reaches a fixed group size, the client compresses
the accumulated   chunk group, adds a chunk group index  to the beginning of the group, and then
appends the header and data  to the corresponding VM file.
A new container  index entry is also created for each chunk group and is written to the corresponding
container index file.

\item
\noindent\textbf{Get()}.
Fetching PDS data is straightforward since each reference contains 
the file offset, and the size of a PDS chunk is available from a segment recipe.
We also maintain a small data cache for the PDS data service to speedup common data fetching.
To read a non-PDS chunk using its reference with container ID and local chunk ID,  the snapshot store client 
first loads the
corresponding VM's container index file specified by the container ID, then searches the chunk
groups  using their  chunk ID coverage.
After that, it reads the identified chunk group from DFS, decompresses it, and seeks to the exact chunk data 
specified by the chunk ID. 
Finally, the client updates its internal chunk cache with the newly loaded content to 
anticipate future sequential reads.

\item
\noindent\textbf{Delete()}.
Chunk deletion occurs when a snapshot expires or gets deleted explicitly by a user. 
%(the overall deletion strategy was discussed in detail in Section~\ref{sect:delete}).
When deletion requests are issued for a specific container,
those requests are simply recorded into the  container's deletion log initially and thus  a lazy
deletion strategy is exercised.
Once local chunk IDs appear in
the deletion log, they will not be referenced by any future snapshot and can be safely deleted when needed. 
This is ensured because we only dedup against the direct parent of a snapshot, so the deleted snapshot's blocks
will only be used if they also exist in other snapshots.
Periodically, the snapshot  store identifies those containers with an excessive
number of deletion requests to  compact and  reclaim the corresponding disk space. 
During compaction, the snapshot store creates a new container (with the same container ID) to replace the 
existing one. This is done by sequentially scanning the old container, copying all the chunks that are not 
found in the deletion log to the new container, and creating new chunk groups and indices. 
Every local chunk ID however is directly copied rather than re-generated. This
process leaves holes in the chunk ID values, but preserves the order and IDs of chunks.
As a result, all data references stored 
in recipes are permanent and stable, and the data reading process
is as efficient as before. Maintaining the stability of chunk IDs also ensures that recipes do not
depend directly on physical storage locations, which simplifies data migration.
\end{itemize}

>>>>

We use the virtual device driver in Xen that employs a bitmap to track the changes 
that have been made to the virtual disk.
Every bit in the bitmap represents a fixed-sized (2MB) segment, indicating whether the segment
has been modified since last backup. 
Segments are further divided into chunks 
%(average 4KB) 
using a content-based chunking algorithm~\cite{frame05}, 
which brings the opportunity of fine-grained deduplication.
When the VM issues a disk write, the dirty bit for the corresponding segment is set
and this indicates such a segments needs to be checked during snapshot backup. 
After the snapshot backup is finished, the driver resets the dirty bit map to a clean state.
For data modification during backup, copy-on-write protection is set so that backup can continue to
copy  a specific version while new changes are recorded.
%copies a frozen version while the modification can still carry on.

The representation of each snapshot has  a two-level index data structure.
The snapshot meta data (called snapshot recipe) contains a list of segments, each of which contains segment
metadata of its chunks (called segment recipe).
In snapshot and segment recipes, 
the data structures  include reference IDs to the actual data location to eliminate the need for additional indirection.

@@

We use the QFS distributed file system to hold snapshot backups.
Following the VM-centric idea for the purpose of fault isolation,
each VM has its own snapshot store, containing new data chunks which are considered
to be non-duplicates.
As shown in Figure~\ref{fig:as_arch}, we explain the data structure of the snapshot stores as follows.
%\begin{itemize}
%\item
There is an independent store containing all PDS chunks shared among different VMs as
a single file.
%All PDS chunks are stored in one PDS file. 
Each reference ID to a PDS data chunk in the PDS index is the offset within the PDS file.
%There is an option to use a data structure similar to the non-PDS store. We opt
%for the simpler format because
Additional compression is not applied because 
for the data sets we have tested, we only observed limited spatial locality 
among popular data chunks. On average the number of consecutive PDS index hits is lower than 7,
thus it is not very effective to group a large number of chunks as a compression and data fetch unit. 
For the same reason, we decide not to take the sampled index approach~\cite{Guo2011} 
for detecting duplicates from PDS as limited spatial locality is not sufficient to enable
effective prefetching for sampled indexing.
%MSST-CHANGE-BEGIN
It should be noted that the above spatial locality number for PDS data 
is subject to application characteristics and we plan to study further on this issue in the future.
%MSST-CHANGE-END

@@

For non PDS data, the snapshot store of a VM is  divided into a set of containers and 
each container is approximately 1GB.  The reason for this  division is to simplify the compaction process
conducted periodically. When  chunks are deleted from old snapshots,
chunks without any reference from other snapshots can be removed by this compaction process.
By limiting the size of a container, we can control the length of each round of compaction.
The compaction  can work on one container at a time and move the in-use data chunks to another container. 
%MSST-CHANGE-BEGIN
The 1GB setting is based on our experience that compaction 
performs well and can  complete  in a reasonable amount time.
%MSST-CHANGE-END

Each non-PDS container is further divided into a set of chunk data groups. Each chunk group is composed of
a set of data chunks and is the basic unit in data access and retrieval. 
In writing a chunk during backup, the system accumulates data chunks and stores the entire
group as a unit after compression. This  compression can reduce data by several times  in our tested data.
%MSST-CHANGE-BEGIN
For the experiments, we use LZO2 which is a lossless compression libary and does not affect snapshot
data availability. 
%MSST-CHANGE-END
When accessing a particular chunk, its chunk group is retrieved from the storage
and decompressed. Given the high spatial locality and usefulness of prefetching  in 
snapshot chunk accessing~\cite{Guo2011,Rhea2008},
retrieval of  a data chunk  group naturally works well with prefetching. 
%MSST-CHANGE-BEGIN
A  typical chunk group contains 1000 chunks in our experiment and this setting is effective
for VM snapshot backup we are studying here.
% with an average size of 200-600 chunks.
%MSST-CHANGE-END
%\item
Each non-PDS data container is represented by three files:
%in the DFS
1) the container data file holds the actual content, 
2) the container index file is responsible for translating a data reference
into its location within a container, and 
3) a chunk deletion log file records all the deletion requests within  the container.

%\item
A non-PDS data chunk reference ID stored in the index of snapshot recipes
is composed of two parts: a container ID with 2 bytes and a local chunk ID with 6 bytes.
Each container maintains a local  chunk counter and assigns the current number 
as a chunk ID  when  a new chunk is added to this  container. 
Since data chunks are always appended to a snapshot store during backup, 
local chunk IDs are monotonically increasing.
When a chunk is to be accessed, the segment recipe contains a reference ID pointing to  
a data chunk, which is used to lookup up the chunk data as described shortly.

@@

\item
\noindent\textbf{Append()}. 
For PDS data, the chunk is appended to the end of the PDS file and the offset is returned as the  reference.
This operation  may only be used during PDS recalculation.
For non-PDS data, this call places a chunk into 
the snapshot store and returns a reference ID to be stored in the recipe of a snapshot. 
The write requests to append chunks to a VM store are accumulated at the client side. 
When the number of write requests reaches a fixed group size, the client compresses
the accumulated   chunk group, adds a chunk group index  to the beginning of the group, and then
appends the header and data  to the corresponding VM file.
A new container  index entry is also created for each chunk group and is written to the corresponding
container index file.

\item
\noindent\textbf{Get()}.
Fetching PDS data is straightforward since each reference ID contains 
the file offset, and the size of a PDS chunk is available from a segment recipe.
We also maintain a small data cache for the PDS data service to speedup common data fetching.
To read a non-PDS chunk using its reference ID with container ID and local chunk ID,  the snapshot store client 
first loads the
corresponding VM's container index file specified by the container ID, then searches the chunk
groups  using their  chunk ID coverage.
After that, it reads the identified chunk group from DFS, decompresses it, and seeks to the exact chunk data 
specified by the chunk ID. 
Finally, the client updates its internal chunk cache with the newly loaded content to 
anticipate future sequential reads.

\item
\noindent\textbf{Delete()}.
Chunk deletion occurs when a snapshot expires or gets deleted explicitly by a user. 
%(the overall deletion strategy was discussed in detail in Section~\ref{sect:delete}).
When deletion requests are issued for a specific container,
those requests are simply recorded into the  container's deletion log initially and thus  a lazy
deletion strategy is exercised.
Once local chunk IDs appear in
the deletion log, they will not be referenced by any future snapshot and can be safely deleted when needed. 
This is ensured because we only dedup against the direct parent of a snapshot, so the deleted snapshot's blocks
will only be used if they also exist in other snapshots.
Periodically, the snapshot  store identifies those containers with an excessive
number of deletion requests to  compact and  reclaim the corresponding disk space. 
During compaction, the snapshot store creates a new container (with the same container ID) to replace the 
existing one. This is done by sequentially scanning the old container, copying all the chunks that are not 
found in the deletion log to the new container, and creating new chunk groups and indices. 
Every local chunk ID however is directly copied rather than re-generated. This
process leaves holes in the chunk ID values, but preserves the order and IDs of chunks.
As a result, all data reference IDs stored 
in recipes are permanent and stable, and the data reading process
is as efficient as before. Maintaining the stability of chunk IDs also ensures that recipes do not
depend directly on physical storage locations, which simplifies data migration.

%MSST-CHANGE-BEGIN
It should be emphasized that the deleted chunks are marked in the log but do not get deleted until compaction
time.  Only compaction process will read and use this information.  
New snapshots should not refer a chunk marked as deletion
and referenced data will never enter the delete log. This is guaranteed 
by the no-false-negative property of the bloom filter.
%MSST-CHANGE-END
\end{itemize}

### 2015-05-07 snapshot_system/mass_background.tex
change for msst

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  In a simple implementation,
it is expensive to compare a large number of 
fingerprints
so a number of techniques have been proposed to improve duplicate identification. 
For example, the data domain method ~\cite{bottleneck08} 
uses an in-memory Bloom filter to identify non-duplicates and a prefetching cache for potential
duplicates that might hit in the future.
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
%Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%to reduce memory requirements at the cost of some loss of deduplication
%efficiency.
Additional inline deduplication and  approximation techniques
are studied in ~\cite{extreme_binning09,sparseindex09,Srinivasan2012,WeiZhangIEEE}.  
%and a parallel batch solution for cluster-based deduplication is 
%studied in ~\cite{WeiZhangIEEE}. 
% \comments{
% All of the above approaches have focused on optimization of deduplication
% efficiency or speed, and none of them have considered the impact
% of deduplication on fault resilience in a cloud environment.  To do so, we
% examine the following properties of VM snapshot deduplication for converged
% IaaS architectures.
% }
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.

@@

Index sampling with a prefetch cache  is proposed 
in~\cite{Guo2011} for efficient single-machine deduplication. 
This scheme uses 25GB memory per 500TB of raw data
and thus the raw data to memory ratio is 20K:1. The scheme proposed in this paper 
uses three times less memory. We have not incorporated this index 
sampling technique because it is difficult to extend 
for a distributed architecture.  To use a distributed memory version 
of the sampled index, every deduplication
request may access a remote machine for index lookup and the overall overhead 
of access latency for all requests is significant.

@@

While deduplication reduces storage cost, it creates an artificial data dependence among VMs: when a shared data chunk
is lost in the storage, all VMs that refer such chunks after deduplication face a data loss.
The work in ~\cite{Reliability06} identifies such an issue in file systems 
and uses more replication for data shared by more files,
and we will adopt this idea.
When the backup storage is implemented using a distributed file system such as Hadoop and 
Google file system~\cite{googlefs03},  
the file system block  size is often chosen to be large (e.g. 64MB). On the other hand, 
deduplication implementations~\cite{Guo2011,extreme_binning09,bottleneck08,Jin2009,Dong2011}
 typically use smaller chunk sizes (e.g. 4K bytes).
Thus we need to address  the size gap between file system blocks and chunks.  
%in adding more replications for popular chunks.

>>>>

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  
%MSST-CHANGE-BEGIN
Source-side deduplication is studied at a chunk or file
level~\cite{Vrable2009,Tan2010} for data backup and can be
integrated  with global deduplication among multiple clients~\cite{Fu2014}.
In this context, the client side deduplication is considered as the source-side
and the client-side computing resource  restriction is not a primary concern.
Our work considers the source-side deduplication in a cloud cluster environment before data is transferred
to a backup storage and the computing resource constraint at the cluster side is a major concern.
%MSST-CHANGE-END

It is expensive to compare a large number of fingerprints
so a number of techniques have been proposed to improve duplicate identification. 
For example, the Data Domain method~\cite{bottleneck08} 
uses an in-memory Bloom filter to identify non-duplicates and a prefetching cache for potential
duplicates that might hit in the future.
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
%Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%to reduce memory requirements at the cost of some loss of deduplication
%efficiency.
Additional inline deduplication and  approximation techniques
are studied in the previous work~\cite{extreme_binning09,sparseindex09,Srinivasan2012,WeiZhangIEEE}.  
%and a parallel batch solution for cluster-based deduplication is 
%studied in ~\cite{WeiZhangIEEE}. 
% \comments{
% All of the above approaches have focused on optimization of deduplication
% efficiency or speed, and none of them have considered the impact
% of deduplication on fault resilience in a cloud environment.  To do so, we
% examine the following properties of VM snapshot deduplication for converged
% IaaS architectures.
% }
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.

@@

Index sampling with a prefetch cache~\cite{Guo2011} 
is proposed 
for efficient single-machine deduplication. 
This scheme uses 25GB memory per 500TB of raw data
and thus the raw data to memory ratio is 20K:1. The scheme proposed in this paper 
uses three times less memory. We have not incorporated this index 
sampling technique because it is difficult to extend 
for a distributed architecture.  To use a distributed memory version 
of the sampled index, every deduplication
request may access a remote machine for index lookup and the overall overhead 
of access latency for all requests is significant.

@@

While deduplication reduces storage cost, it creates an artificial data dependence among VMs: when a shared data chunk
is lost in the storage, all VMs that refer such chunks after deduplication face a data loss.
Such an issue has been identified in file systems~\cite{Reliability06}
and more replication is used for shared data. we follow this idea and study VM-centric strategies.
When the backup storage is implemented using a distributed file system such as Hadoop and 
Google file system~\cite{googlefs03},  
the file system block  size is often chosen to be large (e.g. 64MB). On the other hand, 
deduplication implementations~\cite{Guo2011,extreme_binning09,bottleneck08,Jin2009,Dong2011}
 typically use smaller chunk sizes (e.g. 4K bytes).
Thus we need to address  the size gap between file system blocks and chunks.  
%in adding more replications for popular chunks.

### 2015-05-07 snapshot_system/mass_dedup.tex
change for msst

Since the previous work typically uses a non-uniform chunk size 
with an average of 4KB or 8KB for the best deduplication 
effectiveness~\cite{Guo2011,extreme_binning09,bottleneck08,Dong2011},
we conduct additional local similarity guided deduplication 
by comparing chunk fingerprints of a dirty segment in a snapshot 
with those in  {\em potentially similar} segments from its parent snapshot. 
We consider a parent  segment is  potentially similar to the current dirty segment if 1) the parent segment
is at the same offset as the current segment.
2) the signature of these segments is the same.
The signature of a segment is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). 
%Note that this definition of content similarity is an approximation~\cite{resemblance97}.  
This above approach is motivated by the previous work for similarity-based 
deduplication~\cite{shingling97,extreme_binning09,Dong2011,Aronovich2009} while our focus is local search.

>>>>

Since the previous work typically uses a non-uniform chunk size 
with an average of 4KB or 8KB for the best deduplication 
effectiveness~\cite{Guo2011,extreme_binning09,bottleneck08,Dong2011},
we conduct additional local similarity guided deduplication 
by comparing chunk fingerprints of a dirty segment in a snapshot 
with those in  {\em potentially similar} segments from its parent snapshot.

%MSST-CHANGE-BEGIN
Similarity-guided search is effective for cases that  data chunks are deduplicated irregularly.
For example, we found from the Alibaba's cloud cluster dataset that 
data segment movement happens frequently and  files are  often rewritten rather than being
modified in place.  A dirty-bit or offset-based detection is not able to detect such a movement.
%MSST-CHANGE-END

We consider a parent  segment is  potentially similar to the current dirty segment if 1) the parent segment
is at the same offset as the current segment.
2) the signature of these segments is the same.
The signature of a segment is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). 
%Note that this definition of content similarity is an approximation~\cite{resemblance97}.  
This above approach is motivated by the previous work for similarity-based 
deduplication~\cite{shingling97,extreme_binning09,Dong2011,Aronovich2009} while our focus is local search.
